---
title: "Zombies Homework in Module 8"
author: "Zeynep Senveli"
date: "10/1/2017"
output: html_document
---


Reading in the Zombies.csv file:
```{r}
library(ggplot2)
library(curl)
f <- f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall17/zombies.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```

Calculating the population mean for different variables:
```{r}
mean.height <- mean(d$height)
mean.height
mean.weight <- mean(d$weight)
mean.weight
mean.zk <- mean(d$zombies_killed)
mean.zk
mean.edu <- mean(d$years_of_education)
mean.edu
mean.age <- mean(d$age)
mean.age
```

Since the embedded function uses the sample degrees of freedom (n-1), I need to write up a new function to calculate the standard deviation for the population:
```{r}
stadev <- function(x) {
standard.deviation <- sqrt(sum((x - mean(x))^2)/(length(d$id)))
return(standard.deviation)
}
```

Calculating the population standard deviation for the variables:
```{r}
stdev.height <- stadev(d$height)
stdev.height
stdev.weight <- stadev(d$weight)
stdev.weight
stdev.zk <- stadev(d$zombies_killed)
stdev.zk
stdev.edu <- stadev(d$years_of_education)
stdev.edu
stdev.age <- stadev(d$age)
stdev.age 
```

Plotting the variables by gender using ggplot2:
```{r}
library(ggplot2)
height.boxplot <- ggplot(data = d, aes(x = gender, y = height))
height.boxplot <- height.boxplot + geom_boxplot()
height.boxplot <- height.boxplot + theme(axis.text.x = element_text(angle = 90))
height.boxplot <- height.boxplot + ylab("Height")
height.boxplot
weight.boxplot <- ggplot(data = d, aes(x = gender, y = weight))
weight.boxplot <- weight.boxplot + geom_boxplot()
weight.boxplot <- weight.boxplot + theme(axis.text.x = element_text(angle = 90))
weight.boxplot <- weight.boxplot + ylab("Weight")
weight.boxplot
zk.boxplot <- ggplot(data = d, aes(x = gender, y = zombies_killed))
zk.boxplot <- zk.boxplot + geom_boxplot()
zk.boxplot <- zk.boxplot + theme(axis.text.x = element_text(angle = 90))
zk.boxplot <- zk.boxplot + ylab("Zombies Killed")
zk.boxplot
edu.boxplot <- ggplot(data = d, aes(x = gender, y = years_of_education))
edu.boxplot <- edu.boxplot + geom_boxplot()
edu.boxplot <- edu.boxplot + theme(axis.text.x = element_text(angle = 90))
edu.boxplot <- edu.boxplot + ylab("Years of Education")
edu.boxplot
age.boxplot <- ggplot(data = d, aes(x = gender, y = age))
age.boxplot <- age.boxplot + geom_boxplot()
age.boxplot <- age.boxplot + theme(axis.text.x = element_text(angle = 90))
age.boxplot <- age.boxplot + ylab("Age")
age.boxplot
```

Making scatter plots for height and weight by age:
```{r}
height.scplot <- ggplot(data = d, aes(x = age, y = height))
height.scplot <- height.scplot + xlab("Age") + ylab("Height")  
height.scplot <- height.scplot + geom_point() 
height.scplot <- height.scplot + theme(legend.position = "bottom", legend.title = element_blank()) 
height.scplot
weight.scplot <- ggplot(data = d, aes(x = age, y = weight))
weight.scplot <- weight.scplot + xlab("Age") + ylab("Weight")  
weight.scplot <- weight.scplot + geom_point() 
weight.scplot <- weight.scplot + theme(legend.position = "bottom", legend.title = element_blank()) 
weight.scplot
```
Plotting histograms to see if the quantitative variables have been drawn from a normally distributed population:
```{r}
par(mfrow = c(2, 3))  
attach(d)
hist(height, freq = FALSE, col = "tomato4", main = "Height Distribution", xlab = "Height", 
    ylab = "Density")
hist(weight, freq = FALSE, col = "cyan4", main = "Weight Distribution", xlab = "Weight", 
    ylab = "Density")
hist(age, freq = FALSE, col = "olivedrab", main = "Age Distribution", xlab = "Age", 
    ylab = "Density")
hist(zombies_killed, freq = FALSE, col = "coral", main = "Killed Zombie Distribution", xlab = "Zombies Killed",     ylab = "Density")
hist(years_of_education, freq = FALSE, col = "lemonchiffon4", main = "Distribution of Years of Education", xlab =       "Years of Education", ylab = "Density")
detach(d)
```
According to the histograms, height, weight, and age seem to be drawn from normally distributions when zombies killed and years of education are skewed to the right (positively skewed) so possibly a Poisson distribution. 


Sampling (once) a set of 30 survivors and calculating the sample's mean and standard deviation:
```{r}
s <- d[sample(nrow(d), 30), ]
s
```

Means of the quantitative variables in my sample s of size 30:
```{r}
mean.s.height <- mean(s$height)
mean.s.height
mean.s.weight <- mean(s$weight)
mean.s.weight
mean.s.age <- mean(s$age)
mean.s.age
mean.s.zk <- mean(s$zombies_killed)
mean.s.zk
mean.s.edu <- mean(s$years_of_education)
mean.s.edu
```
Standard deviations of the quantitative variables in my sample s of size 30:
```{r}
sd.s.height <- sd(s$height)
sd.s.height
sd.s.weight <- sd(s$weight)
sd.s.weight 
sd.s.age <- sd(s$age)
sd.s.age
sd.s.zk <- sd(s$zombies_killed)
sd.s.zk 
sd.s.edu <- sd(s$years_of_education)
sd.s.edu
```
Calculating the standard error of the sample s:
```{r}
se.s.height <- sd.s.height/sqrt(30)
se.s.height
se.s.weight <- sd.s.weight/sqrt(30)
se.s.weight
se.s.age <- sd.s.age/sqrt(30)
se.s.age
se.s.zk <- sd.s.zk/sqrt(30)
se.s.zk
se.s.edu <- sd.s.edu/sqrt(30)
se.s.edu
```

Building confidence intervals of 95% for the quantitative variables for the sample s:
```{r}
lower.CI.s.height <- mean(s$height)-(qt(p=0.975,df=29)*se.s.height)
upper.CI.s.height <- mean(s$height)+(qt(p=0.975,df=29)*se.s.height)
CI.s.height <- c(lower.CI.s.height, upper.CI.s.height)
CI.s.height 
lower.CI.s.weight <- mean(s$weight)-(qt(p=0.975,df=29)*se.s.weight)
upper.CI.s.weight <- mean(s$weight)+(qt(p=0.975,df=29)*se.s.weight)
CI.s.weight <- c(lower.CI.s.weight, upper.CI.s.weight)
CI.s.weight
lower.CI.s.age <- mean(s$age)-(qt(p=0.975,df=29)*se.s.age)
upper.CI.s.age <- mean(s$height)+(qt(p=0.975,df=29)*se.s.age)
CI.s.age <- c(lower.CI.s.age, upper.CI.s.age)
CI.s.age
lower.CI.s.zk <- mean(s$zombies_killed)-(qpois(p=0.975,lambda=mean(s$zombies_killed))*se.s.zk)
upper.CI.s.zk <- mean(s$zombies_killed)+(qpois(p=0.975,lambda=mean(s$zombies_killed))*se.s.zk)
CI.s.zk <- c(lower.CI.s.zk, upper.CI.s.zk)
CI.s.zk #Confidence Inteval - Sample - Zombies killed
lower.CI.s.edu <- mean(s$years_of_education)-(qpois(p=0.975,lambda=mean(s$years_of_education))*se.s.edu)
upper.CI.s.edu <- mean(s$years_of_education)+(qpois(p=0.975,lambda=mean(s$years_of_education))*se.s.edu)
CI.s.edu <- c(lower.CI.s.edu, upper.CI.s.edu)
CI.s.edu #Confidence Inteval - Sample - Years of Education
```

Drawing 99 samples (each of size 30) from zombie survivor population:

#height and mean
```{r}
k <- 99  
ht <- NULL
for (i in 1:k) {
    ht[[i]] <- sample(s$height)
}
head(ht)

mh <- NULL
for (i in 1:k) {
    mh[i] <- mean(ht[[i]])
}
mean(mh)
```
#weight and mean
```{r}
k <- 99  
wt <- NULL
for (i in 1:k) {
    wt[[i]] <- sample(s$weight)
}
head(wt)

mw <- NULL
for (i in 1:k) {
    mw[i] <- mean(wt[[i]])
}
mean(mw)
```

#age and mean
```{r}
k <- 99  
at <- NULL
for (i in 1:k) {
    at[[i]] <- sample(s$age)
}
head(at)

ma <- NULL
for (i in 1:k) {
    ma[i] <- mean(at[[i]])
}
mean(ma)
```

#zombies killed and mean
```{r}
k <- 99  
zt <- NULL
for (i in 1:k) {
    zt[[i]] <- sample(s$zombies_killed)
}
head(zt)

mz <- NULL
for (i in 1:k) {
    mz[i] <- mean(zt[[i]])
}
mean(mz)
```

#years of education and mean
```{r}
k <- 99  
et <- NULL
for (i in 1:k) {
    et[[i]] <- sample(s$years_of_education)
}
head(et)

me <- NULL
for (i in 1:k) {
    me[i] <- mean(et[[i]])
}
mean(me)
```

